---
title: "DR Test"
author: "Brian Richardson"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output:
  html_document:
    code_folding: hide
---

```{r message = F, warning = F}

rm(list = ls())
library(dplyr)
library(ggplot2)
library(MASS)
library(tmvtnorm)
library(devtools)
library(truncnorm)
load_all()

```

The goal of this program is to develop a simulation setting for the DR estimator.

### Scenario

* Outcome model: $\textrm{E}(Y|A,\pmb{L})=0.35 + 0.15A + 0.25L_1 + 0.2L_2 + 0.05AL_1 + 0.1AL_2$
* MSM: $\textrm{E}\{Y(a)\} = 0.35 + 0.25(0.5) + \{0.15 + 0.05(0.5)\}a = 0.475 + 0.175a$


```{r}

n = 2000; vare = 0.015; B = 80; seed = 1;

# define parameters -------------------------------------------------------

cov.e = vare
mc.seed <- 123                                # MCCS seed
inv.link <- inv.ident                         # inverse link
d.inv.link <- d.inv.ident                     # deriv of inv link
g <- c(0.35, 0.15, 0.25, 0.2, 0.05, 0.1)      # outcome model parameters
formula <- "~A*L1 + A*L2"                     # outcome model formula
ps.formula <- "~L1 + L2"                      # propensity score model formula
ipw.formula <- "~A"                           # ipw.formula
formula.inc <- "~A*L2"                        # incorrect outcome model
ps.formula.inc <- "~L2"                       # incorrect propensity score model

# simulate data -----------------------------------------------------------

set.seed(seed)
L1 <- rbinom(n, 1, 0.5)                                  # confounder 1
L2 <- rnorm(n, 0, 0.16)                                     # confounder 2
EA <- 0.1 - 0.2*L1 + 0.6*L2
A <- rnorm(n, EA, sqrt(0.04))          # exposure
Astar <- A + rnorm(n, 0, sqrt(cov.e))
a <- c(0, 1)                     # grid of exposures
#apply(mm, 2, range)
#0.25 / apply(mm, 2, function(x) max(x) - min(x))
EY <- inv.link(model.matrix(as.formula(formula)) %*% g)  # mean of outcome
#plot(L1, A); plot(L2, A); plot(A, EY); plot(L1, EY); plot(L2, EY);
hist(EY); range(EY)
EY[EY < 0] <- 0; EY[EY > 1] <- 1
Y <- rbinom(n, 1, EY)                                    # outcome
dat0 <- data.frame(Y, A, L1, L2)                         # oracle data
datstar <- data.frame(Y, A = Astar, L1, L2)              # measured data
args <- list(formula = formula,                          # arguments for fitting
             ps.formula = ps.formula,
             inv.link = inv.link,
             d.inv.link = d.inv.link)

```

```{r}

dr.res0 <- fit.dr(data = dat0,
                 args = args,
                 a = a)

round(dr.res0$est, 4)
c(0.475, 0.475 + 0.175) # truth


```

```{r}

dr.res <- fit.dr.mccs(data = datstar,
                      args = args,
                      a = a,
                      cov.e = cov.e,
                      B = B,
                      mc.seed = mc.seed)

round(dr.res$est, 4)
c(0.475, 0.475 + 0.175) # truth


```


```{r}

# compute estimates and std errors of MSM coefficient for a ---------------

get.est.se.a <- function(res, res.list) {
  name <- strsplit(res, "[.]")[[1]]

  # for IPW, extract estimate and std error for coefficient of a
  if (name[1] == "ipw") {
    est <- unname(res.list[[res]]$est[2])
    se <- sqrt(diag(res.list[[res]]$var)[2])
    bse <- sqrt(diag(res.list[[res]]$bc.var)[2])

    # for g-fmla and double robust, use delta method on E{Y(1)}, E{Y(0)}
  } else {
    est <- diff(tail(unname(res.list[[res]]$est), 2))
    vec <- numeric(length(res.list[[res]]$est))
    vec[(length(vec)-1):length(vec)] <- c(1, -1)
    se <- sqrt(vec %*% res.list[[res]]$var %*% vec)
    bse <- sqrt(vec %*% res.list[[res]]$bc.var %*% vec)
  }
  c(method = name[1],
    type = name[2],
    est = est,
    se = se,
    bse = bse)
}

# (oracle) x (gfmla, ipw, dr) estimates given model specs
est.all <- function(ps.formula, formula) {

  # store results in list
  res.list = list()

  # length of outcome model params
  len.out <- ncol(model.matrix(as.formula(formula), data = dat0))

  # g-formula
  gfmla.args <- list(formula = formula,
                     inv.link = inv.link,
                     d.inv.link = d.inv.link)
  res.list[["gfmla.oracle"]] <- fit.gfmla(
    data = dat0, args = gfmla.args, a = c(0, 1))

  # ipw
  ipw.args <- list(formula = ipw.formula,
                   ps.formula = ps.formula,
                   inv.link = inv.link,
                   d.inv.link = d.inv.link)
  res.list[["ipw.oracle"]] <- fit.ipw(
    data = dat0, args = ipw.args)

  # dr
  dr.args <- list(formula = formula,
                  ps.formula = ps.formula,
                  inv.link = inv.link,
                  d.inv.link = d.inv.link)
  res.list[["dr.oracle"]] <- fit.dr(
    data = dat0, args = dr.args, a = c(0, 1))

  dat <- as.data.frame(t(vapply(
    X = names(res.list),
    FUN = function(res) get.est.se.a(res = res, res.list = res.list),
    FUN.VALUE = character(5)))) |>
    mutate_at(c("est", "se", "bse"), as.numeric)
  
  
    return(dat)
}


# combine results ---------------------------------------------------------

# (00) both models correct
res.00 <- est.all(ps.formula = ps.formula,
                  formula = formula)

# (10) PS incorrect, outcome correct
res.10 <- est.all(ps.formula = ps.formula.inc,
                  formula = formula)

# (01) PS correct, outcome incorrect
res.01 <- est.all(ps.formula = ps.formula,
                  formula = formula.inc)

# (11) both models incorrect
res.11 <- est.all(ps.formula = ps.formula.inc,
                  formula = formula.inc)

# combine results (12 x 11 data frame)
res <- cbind(n = n, B = B, vare = vare, seed = seed,
             rbind(cbind(ps = 0, out = 0, res.00),
                   cbind(ps = 1, out = 0, res.10),
                   cbind(ps = 0, out = 1, res.01),
                   cbind(ps = 1, out = 1, res.11)))

  
```


```{r}

plot(L, A[,1])
plot(L, A[,2])

# plot observed (A1, A2) by L quantile
dat2 %>% 
  mutate(L.grp = factor(ntile(L, 9))) %>% 
  ggplot(aes(x = A1, y = A2)) +
  geom_point(alpha = 0.05) +
  facet_wrap(~L.grp) +
  labs(x = "A1",
       y = "A2") +
  geom_hline(yintercept = 0,
             color = "blue") +
  geom_vline(xintercept = 0,
             color = "blue")

```


We can fit this using the g-formula method and get essentially the correct MSM.

```{r}

# Fit using g-formula -----------------------------------------------------
args <- list(formula = "~A1 + A2 + L",
             inv.link = inv.ident,
             d.inv.link = d.inv.ident)

res.G2 <- fit.gfmla(
  data = dat2,
  a = matrix(c(0, 0,
               1, 0,
               0, 1),
             nrow = 3, ncol = 2,
             byrow = T),
  args = args, return.var = F, return.bcvar = F
)

res.G2$est
EYa <- unname(res.G2$est[5:7])
c(intercept = EYa[1],
  a1 = EYa[2] - EYa[1],
  a2 = EYa[3] - EYa[1])

```

However, when we use IPW here, we get an estimate far from the true MSM parameters (FIXED WITH ADJUSTED A|L PARAMS).

```{r}

# estimate MSM parameters -------------------------------------------------

mean.a <- colMeans(dat2[, c("A1", "A2")])
cov.a <- cov(dat2[, c("A1", "A2")])

#wts.denom <- dmvnorm(dat2[, c("A1", "A2")] - mean.a.l,
#                     mean = c(0, 0),
#                     sigma = var.a.l)
#wts.num <- dmvnorm(dat2[, c("A1", "A2")],
#                   mean = mean.a,
#                   sigma = cov.a)
#wts <- wts.num / wts.denom

wts <- get.SW(
  data = dat2,
  ps.formula = "~L + I(L^2)",
  coef.a.l = t(coef.a.l),
  var.a.l = diag(var.a.l),
  mean.a = mean.a,
  cov.a = cov.a
)

res.I2 <- lm(Y ~ A1 + A2,
             dat2,
             weights = wts)
res.I2$coefficients

```

### Positivity Issue?

Why do we get this issue with IPW and not with G-formula? I think so. Reducing the coefficients for $A|L$ seemed to solve the issue.

