---
title: "IPW Test 2"
author: "Brian Richardson"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output:
  html_document:
    code_folding: hide
---

```{r message = F, warning = F}

rm(list = ls())
library(dplyr)
library(ggplot2)
library(MASS)
library(tmvtnorm)
library(devtools)
library(truncnorm)
load_all()

```

The goal of this program is to develop a simulation setting for the IPW estimator.

### Scenario

Consider a new scenario with a bivariate exposure:

* confounder $L \sim N(0, 1)$,
* exposure $\pmb{A}=(A_1,A_2)$ with $\pmb{A}|L \sim N_2\left(\begin{bmatrix} 0.5L^2 \\ -0.5L^2 \end{bmatrix}, \begin{bmatrix} 1 & 0 \\ 0 & 1 \end{bmatrix}\right)$
* outcome $Y$ with $Y|\pmb{A},L \sim N(A_1 - A_2 + 2L,1)$.

This results in an MSM of $\textrm{E}\{Y(\pmb{a})\} = a_1 - a_2$. We generate data from the above DGP with a large sample size $n=10000$.

```{r}

# parameters --------------------------------------------------------------

seed <- 1
n <- 10000                            # sample size
gg <- c(0, 1, 1, 1)                   # Y|A,L parameters
coef.a.l <- matrix(c(0, 0,
                     0, 0,
                     1, -1),
                   nrow = 3, byrow = T)
var.a.l <- matrix(c(1, 0,
                    0, 1),
                  nrow = 2)

# simulate data -----------------------------------------------------------

set.seed(seed)                                 # for reproducibility
L <- rnorm(n, 0, 0.36)
mean.a.l <- cbind(1, L, L^2) %*% coef.a.l
A <- rmvnorm(n = n, mean = c(0, 0), sigma = var.a.l) + mean.a.l
Y.mean <- cbind(1, A, L) %*% gg
Y <- rnorm(n, 0, 1) + Y.mean                   # outcome Y
dat2 <- data.frame(Y, A, L)                    # oracle data
colnames(dat2) <- c("Y", "A1", "A2", "L")

```


```{r}

plot(L, A[,1])
plot(L, A[,2])

# plot observed (A1, A2) by L quantile
dat2 %>% 
  mutate(L.grp = factor(ntile(L, 9))) %>% 
  ggplot(aes(x = A1, y = A2)) +
  geom_point(alpha = 0.05) +
  facet_wrap(~L.grp) +
  labs(x = "A1",
       y = "A2") +
  geom_hline(yintercept = 0,
             color = "blue") +
  geom_vline(xintercept = 0,
             color = "blue")

```


We can fit this using the g-formula method and get essentially the correct MSM.

```{r}

# Fit using g-formula -----------------------------------------------------
args <- list(formula = "~A1 + A2 + L",
             inv.link = inv.ident,
             d.inv.link = d.inv.ident)

res.G2 <- fit.gfmla(
  data = dat2,
  a = matrix(c(0, 0,
               1, 0,
               0, 1),
             nrow = 3, ncol = 2,
             byrow = T),
  args = args, return.var = F, return.bcvar = F
)

res.G2$est
EYa <- unname(res.G2$est[5:7])
c(intercept = EYa[1],
  a1 = EYa[2] - EYa[1],
  a2 = EYa[3] - EYa[1])

```

However, when we use IPW here, we get an estimate far from the true MSM parameters (FIXED WITH ADJUSTED A|L PARAMS).

```{r}

# estimate MSM parameters -------------------------------------------------

mean.a <- colMeans(dat2[, c("A1", "A2")])
cov.a <- cov(dat2[, c("A1", "A2")])

#wts.denom <- dmvnorm(dat2[, c("A1", "A2")] - mean.a.l,
#                     mean = c(0, 0),
#                     sigma = var.a.l)
#wts.num <- dmvnorm(dat2[, c("A1", "A2")],
#                   mean = mean.a,
#                   sigma = cov.a)
#wts <- wts.num / wts.denom

wts <- get.SW(
  data = dat2,
  ps.formula = "~L + I(L^2)",
  coef.a.l = t(coef.a.l),
  var.a.l = diag(var.a.l),
  mean.a = mean.a,
  cov.a = cov.a
)

res.I2 <- lm(Y ~ A1 + A2,
             dat2,
             weights = wts)
res.I2$coefficients

```

### Positivity Issue?

Why do we get this issue with IPW and not with G-formula? I think so. Reducing the coefficients for $A|L$ seemed to solve the issue.

