---
title: "MCCS IPW Estimated Measurement Error Covariance Simulation Results"
author: "Brian Richardson"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output:
  html_document:
    code_folding: hide
---

```{r message = F, warning = F}

rm(list = ls())
library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggh4x)
library(kableExtra)

```

The goal of this program is to evaluate the Monte Carlo corrected score inverse probability weighting (MCCS IPW) estimator in the presence of confounding and measurement error, where the measurement error covariance is estimated using supplemental data.

## Setup

Consider the following setup:

* continuous confounder $L \sim U(0, 1)$,
* trivariate continuous exposure $\pmb{A} = (A_1, A_2, A_3)^T$ with $\pmb{A}|L$ having multivariate normal distribution $N_3\left(\begin{bmatrix} 0.4L \\ -0.4L \\ 0.2 - 0.1L \end{bmatrix}, \begin{bmatrix} 0.09 & 0 & 0 \\ 0 & 0.09 & 0 \\ 0 & 0 & 0.09 \end{bmatrix}\right)$. 
* mismeasured exposure $\pmb{A}^* = (A^*_1, A^*_2,  A^*_3)^T = \pmb{A} + \pmb{\epsilon}$, where $\pmb{\epsilon} = (\epsilon_1, \epsilon_2, \epsilon_3)^T \sim N(\pmb{0}, \pmb{\Sigma}_e)$, $\pmb{\Sigma}_e = \begin{bmatrix} \sigma_e^2 & 0 & 0 \\ 0 & \sigma_e^2 & 0 \\ 0 & 0 & 0\end{bmatrix}$.
* binary outcome $Y$ with $\textrm{E}(Y|\pmb{A},L) =  \widetilde\gamma_0 + \widetilde\gamma_1 a_1 + \widetilde\gamma_2 a_2 + \widetilde\gamma_3 a_3 + \widetilde\gamma_4L + \widetilde\gamma_5 a_1L + \widetilde\gamma_6 a_2L + \widetilde\gamma_7 a_3L$, where $\widetilde{\pmb{\gamma}} = (\widetilde\gamma_0, \widetilde\gamma_1, \widetilde\gamma_2, \widetilde\gamma_3, \widetilde\gamma_4, \widetilde\gamma_5, \widetilde\gamma_6, \widetilde\gamma_7)^T = (0.4, 0.15, 0.15, 0.2, 0.1, 0.1, 0, -0.1)^T$.

This data generating process leads to the marginal structural model
$\textrm{E}\{Y(\pmb{a})\} = \gamma_0 + \gamma_1 a_1 + \gamma_2 a_2 + \gamma_3 a_3$, where $\pmb{\gamma} = (\gamma_0, \gamma_1, \gamma_2, \gamma_3)^T = (0.45, 0.2, 0.15, 0.15)^T$.

The estimand of interest in these simulations is the parameter $\pmb{\gamma}$ in the marginal structural model.

For these simulations, the covariance $\pmb{\Sigma}_e$ of the measurement error $\pmb{\epsilon}$ is unknown and estimated using supplemental data. In particular, for $5$ subjects, we observe $k$ replicates of $\pmb{A}^*$, and estimate $\pmb{\Sigma}_e$ with

$$\widehat{\pmb{\Sigma}}_e = \frac{\sum_{i=1}^{5} \sum_{j=1}^{k} (\pmb{A}^*_{ij} - \pmb{A}^*_{i.})^{T}(\pmb{A}^*_{ij} - \pmb{A}^*_{i.})}{\sum_{i=1}^{5}(k - 1)}.$$

## Methods

In these simulations, we only look at the IPW CS estimator. We vary the sample size $n$ and the number of exposure replicates $k$.

## Results

```{r}

# load simulation results from each of 10 clusters
sim.out.list <- lapply(
  X = 0:9,
  FUN = function(clust) {
    cbind(clust,
          read.csv(paste0("sim_data/ipw_estvar_data/sd",
                          clust, ".csv")))
  })


# combine simulation results into 1 data frame
sim.out <- bind_rows(sim.out.list)

# true MSM parameters
g <- c(0.45, 0.20, 0.15, 0.15)

# number of sims per setting
n.rep <- nrow(sim.out) /
  n_distinct(dplyr::select(filter(sim.out, !is.na(n)), n, k))

# make long data frame
sim.out.long <- sim.out %>% 
  pivot_longer(cols = starts_with("ghat") |
                 starts_with("stde") |
                 starts_with("bste"), 
               names_to = "method.param",
               values_to = "val") %>% 
    mutate(method = factor(substr(method.param, 6, 7),
                           levels = c("CI"),
                           labels = c("IPW CS")),
           param = factor(substr(method.param, 9, 9)),
           name = factor(substr(method.param, 1, 4)),
           g.true = g[param]) %>% 
  dplyr::select(-method.param) %>% 
  group_by(clust, n, B, vare, k, method, param, name) %>% 
  mutate(id = row_number()) %>% 
  pivot_wider(names_from = name,
              values_from = val,
              id_cols = c(clust, n, B, vare, k, method,
                          param, g.true, id)) %>% 
  mutate(ci.lower = ghat - qnorm(0.975) * stde,
         ci.upper = ghat + qnorm(0.975) * stde,
         ci.cov = g.true >= ci.lower & g.true <= ci.upper,
         bcci.lower = ghat - qnorm(0.975) * bste,
         bcci.upper = ghat + qnorm(0.975) * bste,
         bcci.cov = g.true >= bcci.lower & g.true <= bcci.upper)

```

We first check for simulations with errors.

```{r}

# summarize proportion of missing data by setting
sim.out.long %>% 
  filter(param == 1) %>% 
  group_by(method, n, k) %>% 
  summarise(prop.error = mean(is.na(ghat)),
            n.error = sum(is.na(ghat))) %>% 
  filter(prop.error > 0) %>% 
  ungroup()

```

### Plots of simulation results

Below are boxplots of the empirical distributions of estimators for $\pmb{\gamma}$.

```{r}

# extract simulation parameters
n <- unique(sim.out$n)
k <- unique(sim.out$k)

# make labels for plots
n.labs <- paste0("n = ", n)
names(n.labs) <- n
k.labs <- paste0("k = ", k)
names(k.labs) <- k

param.labs <- paste0("\u03b3", c("0", "1", "2", "3"))
names(param.labs) <- 1:4

```

```{r}

# create plots
make.plot <- function(title = F, y.width = 3) {
  
  # summarize missing or out of bounds estimates
  not.shown <- sim.out.long %>%
    filter(param == 2) %>%
    group_by(method, n, k) %>% 
    summarise(prop.miss = mean(is.na(ghat)),
              prop.out = mean(abs(ghat - g.true > y.width), na.rm = T),
              prop.either = prop.miss + prop.out) %>% 
    filter(prop.either > 0)

  plot <- sim.out.long %>%
    filter(param == 2,
           !is.na(ghat)) %>% 
  mutate(k = factor(k)) %>% 
  ggplot(aes(x = k, y = ghat, group = k)) +
    geom_boxplot(width = 0.7, outlier.size = 0.5) +
    geom_hline(aes(yintercept = g.true),
               linetype = "dashed",
               color = "orange") +
    facet_grid(~n, 
               scales = "free", 
               labeller = labeller(n = n.labs)) +
    labs(y = "Parameter Estimate",
         x = "Number of Exposure Replicates for 5 Individuals",
         fill = "Method",
         color = "Method") +
    ylim(0.15 + y.width * c(-1, 1)) +
    theme_bw() +
    theme(legend.position = "bottom")

  
  if (title) {
    plot <- plot +
      ggtitle(paste0("Empirical Distribution of Parameter Estimates"),
              subtitle = paste0(n.rep, " simulations per setting"))

  }
  return(list(plot = plot,
              tab = not.shown))
}

```

```{r}

# create plots
make.plot.2 <- function(title = F) {
  
  plot <- sim.out.long %>% 
    filter(param == 2,
           !is.na(ghat)) %>%
    pivot_longer(cols = ends_with(".cov"), 
                 names_to = "ci.type",
                 values_to = "ci.cov") %>% 
    mutate(ci.type = ifelse(ci.type == "bcci.cov", "Bias-Corrected", "Uncorrected")) %>% 
    group_by(k, n, ci.type) %>% 
    summarise(cov = mean(ci.cov, na.rm = T)) %>% 
    ggplot(aes(x = k,
               linetype = ci.type,
               y = cov)) +
    geom_line() +
    geom_hline(yintercept = 0.95,
               linetype = "dashed",
               color = "orange") +
    facet_grid(~n,
               scales = "free", 
               labeller = labeller(n = n.labs)) +
    labs(y = "Coverage Probability",
         x = "Number of Exposure Replicates for 5 Individuals",
         linetype = "CI Type") +
    ylim(c(0.5, 1))
    theme_bw() +
    theme(legend.position = "bottom")
  
  if (title) {
    plot <- plot +
      ggtitle(paste0("Empirical Coverage Probabilities of 95% Confidence Intervals"),
              subtitle = paste0(n.rep, " simulations per setting"))

  }
  return(plot)
}

```


### Plots of Simulation Results

Below are boxplots of the empirical distributions of estimators for $\gamma_1$, and plots of coverage probabilities of confidence intervals.

```{r message = F, eval = T}

make.plot()

# save plot
ggsave("sim_figures/ipw_estvar_plot_1.png",
       width = 6,
       height = 4,
       dpi = 300)

```

```{r message = F, eval = T}

make.plot.2()

# save plot
ggsave("sim_figures/ipw_estvar_plot_2.png",
       width = 6,
       height = 4,
       dpi = 300)

```

