---
title: "MCCS G-Formula Simulation Positivity Violation 2 Results"
author: "Brian Richardson"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output:
  html_document:
    code_folding: hide
---

```{r message = F, warning = F}

rm(list = ls())
library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggh4x)
library(devtools)
load_all()

```

The goal of this program is to evaluate the Monte Carlo corrected score G-formula (MCCS G-FMLA) estimator in the presence of confounding and measurement error, with some positivity violation.

## Setup

Consider the following setup:

* two independent binary confounders $L_1 \sim \textrm{Bernoulli}(0.5), L_2 \sim \textrm{Bernoulli}(0.2)$,
* univariate continuous exposure $A$ with $A|L_1,L_2 \sim N(2+0.3L_1-0.5L_2, 0.35)$,
* mismeasured exposure $A^* = A + \epsilon$, where $\epsilon \sim N(0, 0.25)$,
* binary outcome $Y$ with $\textrm{E}(Y|A,L) = \textrm{logit}^{-1}(-2+0.7A-0.6L_1+0.4L_2-0.4AL_1-0.2AL_2)$.

This data generating process leads to a mean potential outcome at $a$ of $\textrm{E}\{Y(a)\} = 0.4\textrm{logit}^{-1}(-2+0.7a) + 0.4\textrm{logit}^{-1}(-2.6+0.3a) + 0.1\textrm{logit}^{-1}(-1.6+0.5a) + 0.1\textrm{logit}^{-1}(-2.2+0.1a)$.

The estimand of interest in these simulations is the dose response curve $\{\textrm{E}[Y(a)] : a \in \mathcal{A}\}$.

## Methods

In these simulations, we compare three methods:

1) Naive G-Formula
2) Corrected GLM
3) Corrected G-Formula

The GLM methods fit a logistic regression mode of $Y$ on $A$ and $L$ with all interaction terms between $A$ and $L$, then estimate $\textrm{E}\{Y(a)\}$ with $\textrm{logit}^{-1}(\widehat\beta_0+\widehat\beta_aa)$. The G-formula methods fit the same logistic regression models, then estimate $\textrm{E}\{Y(a)\}$ with $n^{-1}\sum_{i=1}^n\textrm{logit}^{-1}(\widehat\beta_0+\widehat\beta_Aa+\widehat\beta_{L_1}L_{1i}+\widehat\beta_{L_2}L_{2i}+\widehat\beta_{A:L_1}aL_{1i}+\widehat\beta_{A:L_2}aL_{2i})$.

The oracle methods use the true exposure values $A$, which are not observed in practice. The naive methods treat $A^*$ as $A$, assuming there is no measurement error. The corrected methods use the Monte-Carlo corrected score versions of the estimating functions, which account for measurement error.

## Results

We first check for simulations with errors.

```{r}

# load simulation results from each of 10 clusters
sim.out.list <- lapply(
  X = 0:9,
  FUN = function(clust) {
    cbind(clust,
          read.csv(paste0("sim_data/gfmla_positivity_violation_2_data/sd",
                          clust, ".csv")))
  })


# combine simulation results into 1 data frame
sim.out <- bind_rows(sim.out.list)

# true estimands
g <- c(-2, 0.7, -0.6, 0.4, -0.4, -0.2)          # outcome model parameters
a <- as.numeric(dplyr::select(                  # exposure value of interest
  sim.out, starts_with("a"))[1,]) 
EYa.true <- 0.4 * inv.logit(-2 + 0.7 * a) +     # true dose response curve at a
  0.4 * inv.logit(-2.6 + 0.3 * a) +
  0.1 * inv.logit(-1.6 + 0.5 * a) +
  0.1 * inv.logit(-2.2 + 0.1 * a)

# number of sims per setting
n.rep <- nrow(sim.out) / n_distinct(dplyr::select(sim.out, n, B, vare))

# make long data frame
sim.out.long <- sim.out %>% 
  dplyr::select(!starts_with("a")) %>% 
  pivot_longer(cols = starts_with("OG") | starts_with("NG") | starts_with("CG"), 
               names_to = "method.aa",
               values_to = "val") %>% 
    mutate(method = factor(substr(method.aa, 1, 2),
                           levels = c("OG", "NG", "CG"),
                           labels = c("Oracle G-Formula",
                                      "Naive G-Formula",
                                      "Corrected G-Formula")),
           aa = as.numeric(gsub("[.]", "", substr(method.aa, 4, 5))),
           a = a[aa],
           EYa.true = EYa.true[aa]) %>% 
  dplyr::select(-method.aa)

# make summary data frame
sim.out.summary <- sim.out.long %>% 
  group_by(n, vare, B, method, a) %>% 
  summarise(bias = mean(val - EYa.true),
            se = sd(val) / sqrt(n.rep),
            bias.lower = bias - qnorm(0.975) * se,
            bias.upper = bias + qnorm(0.975) * se)

```

We first check for simulations with errors.

```{r}

# summarize proportion of missing data by setting
sim.out.long %>% 
  filter(aa == 1) %>% 
  group_by(method, n, B, vare) %>% 
  summarise(prop.error = mean(is.na(val))) %>% 
  filter(prop.error > 0) %>% 
  ungroup()

```

```{r}

# colorblind friendly pallette
pal_light <- c('#EE6677', '#228833', '#4477AA', '#CCBB44', '#66CCEE', '#AA3377', '#BBBBBB')
pal_dark <- c('#991122', '#114419', '#223b55', '#6b611d', '#117799', '#55193b', '#5d5d5d')

```

### Plots of Simulation Results

Below is a plot of bias in estimated $\textrm{E}\{Y(a)\}$ over a grid of exposure values.

```{r message = F}

p <- sim.out.summary %>% 
  ggplot(
    aes(x = a,
        y = bias,
        ymin = bias.lower,
        ymax = bias.upper)) +
  geom_point() +
  geom_line() +
  facet_grid(~ method) +
  theme_bw() +
  theme(legend.position = c(0.2, 0.3)) +
  labs(color = "Method",
       shape = "Method",
       fill = "Method",
       y = "Bias")

p

```

```{r}

# save plot
ggsave("sim_figures/gfmla_positivity_violation_sim_plot.png",
       width = 6,
       height = 2.5,
       dpi = 300)

```


