---
title: "Simulation 3: IPW with nonlinear PS model"
author: "Brian Richardson"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output:
  html_document:
    code_folding: hide
---

```{r message = F, warning = F}

rm(list = ls())
library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggh4x)
library(kableExtra)

```

```{r warning = F}

# load simulation results from each of 10 clusters
setwd(dirname(getwd()))
sim.out.list <- lapply(
  X = 0:9,
  FUN = function(clust) {
    cbind(clust,
          read.csv(paste0("sim_data/sim3_ipw_nonlinearps/sd",
                          clust, ".csv")))
  })


# combine simulation results into 1 data frame
sim.out <- bind_rows(sim.out.list)

# true MSM parameters
g <- c(0, 1, 1)

# number of sims per setting
n.rep <- nrow(sim.out) / n_distinct(dplyr::select(sim.out, n, B, vare))

# make long data frame
sim.out.long <- sim.out %>% 
  pivot_longer(cols = starts_with("ghat") |
                 starts_with("stde") |
                 starts_with("bste"), 
               names_to = "method.param",
               values_to = "val") %>% 
    mutate(method = factor(substr(method.param, 6, 7),
                           levels = c("OI", "NI", "RI", "SI", "CI"),
                           labels = c("Oracle IPW",
                                      "Naive IPW",
                                      "Reg. Cal. IPW",
                                      "SIMEX IPW",
                                      "CS IPW")),
           param = factor(substr(method.param, 9, 9)),
           name = factor(substr(method.param, 1, 4)),
           g.true = g[param]) %>% 
  dplyr::select(-method.param) %>% 
  group_by(clust, n, B, vare, method, param, name) %>% 
  mutate(id = row_number()) %>% 
  pivot_wider(names_from = name,
              values_from = val,
              id_cols = c(clust, n, B, vare, method,
                          param, g.true, id)) %>% 
  mutate(ci.lower = ghat - qnorm(0.975) * stde,
         ci.upper = ghat + qnorm(0.975) * stde,
         ci.cov = g.true >= ci.lower & g.true <= ci.upper,
         bcci.lower = ghat - qnorm(0.975) * bste,
         bcci.upper = ghat + qnorm(0.975) * bste,
         bcci.cov = g.true >= bcci.lower & g.true <= bcci.upper)

```

The goal of this program is to evaluate the Monte Carlo corrected score inverse probability weighting (MCCS IPW) estimator in the presence of confounding and measurement error, and in a scenario where the propensity score model $\pmb{A}|L$ is nonlinear.

## Setup

Consider the following setup:

* continuous confounder $L \sim N(0, 0.13)$,
* bivariate continuous exposure $\pmb{A} = (A_1, A_2)^T$ with $\pmb{A}|L$ having multivariate normal distribution $N_2\left(\begin{bmatrix} L^2 \\ -L^2 \end{bmatrix}, \begin{bmatrix} 1 & 0 \\ 0 & 1 \end{bmatrix}\right)$. 
* mismeasured exposure $\pmb{A}^* = (A^*_1, A^*_2)^T = \pmb{A} + \pmb{\epsilon}$, where $\pmb{\epsilon} = (\epsilon_1, \epsilon_2)^T \sim N(\pmb{0}, \pmb{\Sigma}_e)$, $\pmb{\Sigma}_e = \begin{bmatrix} 0.2 & 0 \\ 0 & 0.2 \end{bmatrix}$.
* normal outcome $Y$ with $\textrm{E}(Y|\pmb{A},L) = \widetilde\gamma_0 + \widetilde\gamma_1 A_1 + \widetilde\gamma_2 A_2 + \widetilde\gamma_3 L$, where $\widetilde{\pmb{\gamma}} = (\widetilde\gamma_0, \widetilde\gamma_1, \widetilde\gamma_2, \widetilde\gamma_3)^T = (0, 1, 1, 1)^T$.

This data generating process leads to the marginal structural model
$\textrm{E}\{Y(\pmb{a})\} = \gamma_0 + \gamma_1 a_1 + \gamma_2 a_2$, where $\pmb{\gamma} = (\gamma_0, \gamma_1, \gamma_2)^T = (0, 1, 1)^T$.

The estimand of interest in these simulations is the parameter $\pmb{\gamma}$ in the marginal structural model.

## Methods

In these simulations, we compare five methods:

1) Oracle IPW
2) Naive IPW
3) Regression Calibration IPW
4) SIMEX IPW
5) Corrected IPW

The oracle method uses the true exposure values $\pmb{A}$, which are not observed in practice. The naive method treats $\pmb{A}^*$ as $\pmb{A}$, assuming there is no measurement error. Regression calibration replaces $\pmb{A}^*$ with an approximation of $\textrm{E}(\pmb{A}|\pmb{A}^*,L)$. SIMEX uses simulated data with a sequence of additional measurement error to extrapolate backwards to estimate the dose response curve with no measurement error. The CS method uses the Monte-Carlo corrected score versions of the estimating functions.

## Results

We first check for simulations with errors.

```{r message = F}

# summarize proportion of missing data by setting
sim.out.long %>% 
  filter(param == 1) %>% 
  group_by(method, n, vare, B) %>% 
  summarise(prop.error = mean(is.na(ghat))) %>% 
  filter(prop.error > 0) %>% 
  ungroup()

```

### Plots of simulation results

Below are boxplots of the empirical distributions of estimators for $\pmb{\gamma}$.

```{r}

# extract simulation parameters
n <- unique(sim.out$n)
B <- unique(sim.out$B)
vare <- unique(sim.out$vare)

# make labels for plots
n.labs <- paste0("n = ", n)
names(n.labs) <- n

B.labs <- paste0("B = ", B)
names(B.labs) <- B

vare.labs <- paste0("var(e) = ", vare)
names(vare.labs) <- vare

n.labs <- paste0("n = ", n)
names(n.labs) <- n

param.labs <- paste0("gamma[", c("0", "1", "2"), "]")
names(param.labs) <- 1:3
param.labs <- as_labeller(param.labs,
                          default = label_parsed)

# colorblind friendly pallette
pal_light <- c('#EE6677', '#228833', '#4477AA', '#CCBB44', '#66CCEE', '#AA3377', '#BBBBBB')
pal_dark <- c('#991122', '#114419', '#223b55', '#6b611d', '#117799', '#55193b', '#5d5d5d')

```

```{r}

# create plots
make.plot <- function(title = T, y.lim = Inf, vare. = vare) {
  
  plot <- sim.out.long %>% 
    filter(abs(ghat - g.true) < y.lim,
           vare %in% vare.) %>% 
    ggplot(
    aes(x = method,
        y = ghat,
        fill = method,
        color = method)) +
    geom_boxplot() +
    stat_summary(fun = mean,
                 geom = "point",
                 shape = 8,
                 size = 2,
                 orientation = "x",
                 show.legend = F) +
    geom_hline(aes(yintercept = g.true),
               linetype = "dashed",
               color = "black") +
    facet_grid(param ~ n,
               scales = "free",
               labeller = labeller(n = n.labs,
                                   param = param.labs)) +
    labs(y = "Parameter Estimate",
         fill = "Method",
         color = "Method") +
    theme_bw() +
    theme(axis.title.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.text.x = element_blank(),
          legend.position = "bottom",
          strip.text = element_text()) +
    scale_fill_manual(values = pal_light) +
    scale_color_manual(values = pal_dark)
  
  if (title) {
    plot <- plot + 
      ggtitle(paste0("Empirical Distribution of Parameter Estimates"),
            subtitle = paste0(n.rep, " simulations per setting"))
  }
  return(plot)
}

```

```{r}

make.plot(title = F)

```

```{r warning = F}

# save plot
setwd(dirname(getwd()))
ggsave("sim_figures/sim3_ipw_nonlinearps.png",
       width = 6,
       height = 4,
       dpi = 300)

```


### Table of Simulation Results

Below is a table of the empirical bias, empirical standard error, mean estimated standard error, and empirical coverage probability of the 95% CIs corresponding to $\widehat{\pmb{\gamma}}$, for oracel, naive, and corrected methods, and for $n \in \{400, 800, 8000\}$.

```{r}

tbl <- sim.out.long %>% 
  filter(method %in% c("Oracle IPW", "Naive IPW", "CS IPW")) %>% 
  mutate(Param = paste0("$\\gamma_", as.numeric(param) - 1, "$")) %>% 
  group_by(n, method, Param) %>% 
  summarise(bias = 100 * mean(ghat - g.true),
            emp.se = 100 * sd(ghat),
            est.se = 100 * mean(stde),
            ci.cov = 100 * mean(ci.cov),
            bcest.se = 100 * mean(bste),
            bcci.cov = 100 * mean(bcci.cov))

kbl <- tbl %>% 
  kable(format = "latex",
        digits = c(0, 0, 0,
                   1, 1,
                   1, 1,
                   1, 1),
        align = c(rep("c", 3 + 2),
                  rep("r", 4)),
        booktabs = TRUE,
        linesep = c(rep("", 9), "\\addlinespace"),
        escape = FALSE,
        col.names = c("n", "Method", "Parameter",
                      "Bias", "ESE",
                      "ASE", "Cov",
                      "ASE", "Cov")) %>%
  add_header_above(c(" " = 5 + 2,
                     "UC" = 2,
                     "BC" = 2)) %>% 
  kable_styling("striped") %>% 
  row_spec(row = 0, bold = TRUE)

print(kbl)
  
```


