---
title: "Simulation 4: DR with nonlinear PS model"
author: "Brian Richardson"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output:
  html_document:
    code_folding: hide
---

```{r message = F, warning = F}

rm(list = ls())
library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggh4x)
library(kableExtra)
library(devtools)
load_all()

```

```{r warning = F}

# load simulation results from each of 10 clusters
setwd(dirname(getwd()))
sim.out.list <- lapply(
  X = 0:9,
  FUN = function(clust) {
    cbind(clust,
          read.csv(paste0("sim_data/sim4_dr_nonlinearps/sd",
                          clust, ".csv")))
  })

# true estimand
g <- 0.75

# combine simulation results into 1 data frame
sim.out <- bind_rows(sim.out.list) %>% 
  mutate(g.true = g,
         ci.lower = est - qnorm(0.975) * se,
         ci.upper = est + qnorm(0.975) * se,
         ci.cov = g.true >= ci.lower & g.true <= ci.upper,
         bcci.lower = est - qnorm(0.975) * bse,
         bcci.upper = est + qnorm(0.975) * bse,
         bcci.cov = g.true >= bcci.lower & g.true <= bcci.upper,
         method = factor(method,
                         levels = c("gfmla", "ipw", "dr"),
                         labels = c("G-Formula", "IPW", "DR")),
         type = factor(type,
                       levels = c("oracle", "naive", "rc", "mccs"),
                       labels = c("Oracle", "Naive", "Reg. Cal.", "MCCS")),
         spec = factor(ifelse(ps == 0 & out == 0, "PS and OR",
                       ifelse(ps == 0 & out == 1, "PS",
                       ifelse(ps == 1 & out == 0, "OR", "Neither"))),
                       levels = c("PS and OR", "PS", "OR", "Neither")),
         ps = factor(ps,
                     levels = c(0, 1),
                     labels = c("PS Correct", "PS Incorrect")),
         out = factor(out,
                      levels = c(0, 1),
                      labels = c("Outcome Correct", "Outcome Incorrect")))

# number of sims per setting
n.rep <- nrow(sim.out) /
  n_distinct(dplyr::select(sim.out, n, method, type, ps, out))

```

The goal of this program is to evaluate the Monte Carlo corrected score double robust (MCCS DR) estimator in the presence of confounding, measurement error, and possible model misspecification.

## Setup

Consider the following setup:

* sample size $n\in\{400,2000\}$,
* two independent confounders $L_1 \sim \textrm{Bernoulli}(0.5), L_2 \sim N(1, 0.5)$
* univariate exposure $A$, where $A|L \sim N(2 + 0.9L_1 - 0.6L_2^2, 1.1)$
* continuous outcome $Y$, with $Y|A,L \sim N(1.5 + 0.7A + 0.9L_1 - 0.7L_2 + - 0.6AL_1 + 0.4AL_2, 4)$
* resulting MSM $\textrm{E}\{Y(a)\} = \beta_0 + \beta_1a$, where $\pmb{\beta} = (\beta_0, \beta_1)^T = (1.35, 0.75)^T$
* mismeasured exposure $A^* = A + \epsilon$, where $\epsilon \sim N(0, 0.25)$

The estimand of interest in these simulations is $\beta_1=0.75$, the coefficient for $a$ in the MSM.

## Methods

In these simulations, we consider four scenarios defined by whether the propensity score and outcome models are correctly specified. To misspecify the propensity model, we omit the confounder $L_1$, and to misspecify the outcome model, we omit $L_1$ and its interaction with $A$.

For each scenario, we compare 15 methods based on three estimators:

1) G-Formula
2) IPW
3) DR

and 5 ways to handle measurement error:

1) Oracle
2) Naive
3) Regression Calibration
4) SIMEX
5) MCCS

The oracle method uses the true exposure values $A$, which are not observed in practice. The naive method treats $A^*$ as $A$, assuming there is no measurement error. Regression calibration replaces $A^*$ with an approximation of $\textrm{E}(A|A^*,\pmb{L})$. SIMEX uses simulated data with a sequence of additional measurement error to extrapolate backwards to estimate the dose response curve with no measurement error. The CS method uses the Monte-Carlo corrected score versions of the estimating functions.


## Results

We first check for simulations with errors.

```{r message = F}

# summarize proportion of missing data by setting
sim.out %>% 
  group_by(n, method, type, ps, out) %>% 
  summarise(prop.error = mean(is.na(est))) %>% 
  filter(prop.error > 0) %>% 
  ungroup()

```

```{r}

# colorblind friendly pallette
pal_light <- c('#EE6677', '#228833', '#4477AA', '#CCBB44', '#66CCEE', '#AA3377', '#BBBBBB')
pal_dark <- c('#991122', '#114419', '#223b55', '#6b611d', '#117799', '#55193b', '#5d5d5d')

```

```{r}

# plot all estimators across 4 modeling scenarios
make.plot <- function(n. = 2000, vare. = 0.25) {
  
  sim.out %>% 
    filter(n == n.,
           vare == vare.) %>% 
    ggplot(aes(x = type,
               y = est,
               fill = type,
               color = type)) +
    geom_boxplot() +
    geom_hline(yintercept = g,
               linetype = "dashed",
               color = "orange") +
    facet_nested(method ~ ps + out, scales = "free") +
    labs(y = "Parameter Estimate",
         fill = "Type",
         color = "Type") +
    ggtitle(paste0("Empirical Distribution of Parameter Estimates"),
            subtitle = paste0("n = ", n., "; ",
                              n.rep, " simulations per setting")) +
    theme_bw() +
    theme(axis.title.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.text.x = element_blank()) +
    scale_fill_manual(values = pal_light) +
    scale_color_manual(values = pal_dark)
}

```

### Plot of Simulation Results

Below is a plot of the empirical distributions of the nine estimators under the four model specification scenarios.

```{r message = F}

make.plot(vare. = 0.25)

```

### Table of Simulation Results

The following table shows the empirical bias, empirical standard error (ESE), average estimated standard error (ASE), and empirical coverage probability of the UC and BC 95% CIs for the corrected estimators under the four model specification scenarios and a sample size $n=400$

```{r}

tbl <- sim.out %>% 
  filter(type == "MCCS",
         n == 2000) %>% 
  group_by(spec, method, type) %>% 
  mutate(method = paste0(method, " CS")) %>% 
  summarise(bias = 100 * mean(est - g.true, na.rm = T),
            emp.se = 100 * sd(est, na.rm = T),
            est.se = 100 * mean(se),
            ci.cov = 100 * mean(ci.cov, na.rm = T),
            bcest.se = 100 * mean(bse, na.rm = T),
            bcci.cov = 100 * mean(bcci.cov, na.rm = T))

kbl <- tbl %>% 
  kable(format = "latex",
        align = c(rep("c", 2),
                  rep("r", 6)),
        booktabs = TRUE,
        linesep = c("", "", "\\addlinespace"),
        escape = FALSE,
        digits = c(0, 0,
                   1, 1, 1, 1, 1, 1),
        col.names = c("Correct Specifications", "Method", "Type",
                      "Bias", "ESE", "ASE", "Cov", "ASE", "Cov"),
        caption = "Estimated MSM Slope") %>%
  add_header_above(c(" " = 4,
                     "UC" = 2,
                     "BC" = 2)) %>% 
  kable_styling("striped") %>% 
  row_spec(row = 0, bold = TRUE)

print(kbl)

```

Below is the same table with $n=400$.

```{r}

tbl2 <- sim.out %>% 
  filter(type == "MCCS",
         n == 400) %>% 
  group_by(spec, method, type) %>% 
  mutate(method = paste0(method, " CS")) %>% 
  summarise(bias = 100 * mean(est - g.true, na.rm = T),
            emp.se = 100 * sd(est, na.rm = T),
            est.se = 100 * mean(se),
            ci.cov = 100 * mean(ci.cov, na.rm = T),
            bcest.se = 100 * mean(bse, na.rm = T),
            bcci.cov = 100 * mean(bcci.cov, na.rm = T))

kbl2 <- tbl2 %>% 
  kable(format = "latex",
        align = c(rep("c", 2),
                  rep("r", 6)),
        booktabs = TRUE,
        linesep = c("", "", "\\addlinespace"),
        escape = FALSE,
        digits = c(0, 0,
                   1, 1, 1, 1, 1, 1),
        col.names = c("Correct Specifications", "Method", "Type",
                      "Bias", "ESE", "ASE", "Cov", "ASE", "Cov"),
        caption = "Estimated MSM Slope") %>%
  add_header_above(c(" " = 4,
                     "UC" = 2,
                     "BC" = 2)) %>% 
  kable_styling("striped") %>% 
  row_spec(row = 0, bold = TRUE)

print(kbl2)

```

