---
title: "Simulation 5: g-formula with varying reliability"
author: "Brian Richardson"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output:
  html_document:
    code_folding: hide
---

```{r message = F, warning = F}

rm(list = ls())
library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggh4x)
library(kableExtra)
library(devtools)
load_all()

```

```{r warning = F}

# load simulation results from each of 10 clusters
setwd(dirname(getwd()))
sim.out.list <- lapply(
  X = 0:9,
  FUN = function(clust) {
    cbind(clust,
          read.csv(paste0("sim_data/sim5_gfmla_reliability/sd",
                          clust, ".csv")))
  })

# combine simulation results into 1 data frame
sim.out <- bind_rows(sim.out.list)

# true estimands
g <- c(0, 0.25, 0.5, -0.5, 1)          # outcome model parameters
a <- -1:2                                        # exposure value of interest
EYa.true <- g[1] + 0.5*g[5] + a*g[2] + a^2*g[3] + a^3*g[4]

# number of sims per setting
n.rep <- nrow(sim.out) / n_distinct(dplyr::select(sim.out, n, B, rel))

# make long data frame
sim.out.long <- sim.out %>% 
  pivot_longer(cols = starts_with("est") | 
                      starts_with("ste") |
                      starts_with("bcs"), 
               names_to = "method.aa",
               values_to = "val") %>% 
    mutate(method = factor(substr(method.aa, 5, 6),
                           levels = c("NG", "CG"),
                           labels = c("Naive G-Formula",
                                      "G-Formula CS")),
           aa = factor(substr(method.aa, 8, 8)),
           a = factor(a[aa]),
           name = factor(substr(method.aa, 1, 3)),
           EYa.true = EYa.true[aa]) %>% 
  dplyr::select(-method.aa) %>% 
  group_by(clust, n, B, rel, method, aa, a, name) %>% 
  mutate(id = row_number()) %>% 
  pivot_wider(names_from = name,
              values_from = val,
              id_cols = c(clust, n, B, rel, method, aa, a, EYa.true, id)) %>% 
  mutate(ci.lower = est - qnorm(0.975) * ste,
         ci.upper = est + qnorm(0.975) * ste,
         ci.cov = EYa.true >= ci.lower & EYa.true <= ci.upper,
         bcci.lower = est - qnorm(0.975) * bcs,
         bcci.upper = est + qnorm(0.975) * bcs,
         bcci.cov = EYa.true >= bcci.lower & EYa.true <= bcci.upper)

```


The goal of this program is to evaluate the G-formula CS estimator with a nonlinear outcome model and for a sequence of exposure reliability values.

## Setup

Consider the following setup:

* single confounder $L_1 \sim U(0,1)$,
* univariate continuous exposure $A$ with $A|L \sim N(L,)$,
* mismeasured exposure $A^* = A + \epsilon$, where $\epsilon \sim N(0, 0.09)$,
* normal outcome $Y$ with $Y|A,L \sim N(0 + 0.25A, 0.5A^2, -0.5^2 + L, 0.16)$.

This data generating process leads to a mean potential outcome at $a$ of $\textrm{E}\{Y(a)\} = 0.5 + 0.25a + 0.5a^2 - 0.5a^3$.

The estimand of interest in these simulations is the dose response curve $\{\textrm{E}[Y(a)] : a \in \mathcal{A}\}$.

## Methods

In these simulations, we compare two methods:

1) Naive G-Formula
2) CS G-Formula

The naive method treats $A^*$ as $A$, assuming there is no measurement error. The CS method uses the Monte-Carlo corrected score versions of the estimating functions.

## Results

We first check for simulations with errors.

```{r message = F}

# summarize proportion of missing data by setting
sim.out.long %>% 
  filter(aa == 1) %>% 
  group_by(method, n, B, rel) %>% 
  summarise(prop.error = mean(is.na(est))) %>% 
  filter(prop.error > 0)

```

```{r}

# extract simulation parameters
n <- unique(sim.out$n)

# make labels for plots
method.labs <- c("Naive G-Formula",
                 "G-Formula CS")

names(method.labs) <- c("NG", "CG")

n.labs <- paste0("n = ", n)
names(n.labs) <- n

a.labs <- paste0("a = ", a)
names(a.labs) <- a

rel.lab = unique(round(sim.out.long$rel, 2))
rel.lab[seq(2, n_distinct(rel.lab), 2)] <- ""

# colorblind friendly pallette
pal_light <- c('#EE6677', '#228833', '#4477AA', '#CCBB44', '#66CCEE', '#AA3377', '#BBBBBB')
pal_dark <- c('#991122', '#114419', '#223b55', '#6b611d', '#117799', '#55193b', '#5d5d5d')

```

```{r}

# create plots
make.plot <- function(title = F, a. = 1) {

  plot <- sim.out.long %>%
    filter(aa == a.) %>% 
    mutate(rel = factor(rel)) %>% 
    ggplot(aes(x = rel, y = est, group = rel)) +
    geom_boxplot(width = 0.7, outlier.size = 0.5) +
    geom_hline(aes(yintercept = EYa.true),
               linetype = "dashed",
               color = "orange") +
    facet_grid(n ~ method,
               scales = "free", 
               labeller = labeller(n = n.labs)) +
    labs(y = "Parameter Estimate",
         x = "Exposure Reliability",
         fill = "Method",
         color = "Method") +
    theme_bw() +
    theme(legend.position = "bottom") +
    scale_x_discrete(labels = rel.lab)

  
  if (title) {
    plot <- plot +
      ggtitle(paste0("Empirical Distribution of Parameter Estimates"),
              subtitle = paste0(n.rep, " simulations per setting"))

  }
  return(plot)
}

```


```{r}

# create plots
make.plot.2 <- function(title = F, a. = 1) {
  
  plot <- sim.out.long %>% 
    filter(aa == a.) %>% 
    pivot_longer(cols = ends_with(".cov"), 
                 names_to = "ci.type",
                 values_to = "ci.cov") %>% 
    mutate(ci.type = ifelse(ci.type == "bcci.cov",
                            "Bias-Corrected",
                            "Uncorrected")) %>% 
    group_by(rel, n, method, ci.type) %>% 
    summarise(cov = mean(ci.cov, na.rm = T)) %>% 
    ggplot(aes(x = rel,
               linetype = ci.type,
               y = cov)) +
    geom_line() +
    geom_hline(yintercept = 0.95,
               linetype = "dashed",
               color = "orange") +
    facet_grid(n ~ method,
               scales = "free", 
               labeller = labeller(n = n.labs)) +
    labs(y = "Coverage Probability",
         x = "Exposure Reliability",
         linetype = "CI Type") +
    theme_bw() +
    theme(legend.position = "bottom")
  
  if (title) {
    plot <- plot +
      ggtitle(paste0("Empirical Coverage Probabilities of 95% Confidence Intervals"),
              subtitle = paste0(n.rep, " simulations per setting"))

  }
  return(plot)
}

```

### Plots of Simulation Results

Below are boxplots of the empirical distributions of estimators for $\textrm{E}\{Y(a)\}$ at $a=1$, and plots of coverage probabilities of confidence intervals.

```{r warning = F, message = F, eval = T}

make.plot()

# save plot
setwd(dirname(getwd()))
ggsave("sim_figures/sim5_gfmla_reliability_1.png",
       width = 6,
       height = 4,
       dpi = 300)

```

```{r warning = F, message = F, eval = T}

make.plot.2()

# save plot
setwd(dirname(getwd()))
ggsave("sim_figures/sim5_gfmla_reliability_2.png",
       width = 6,
       height = 4,
       dpi = 300)

```

