---
title: "Simulation 6: IPW with varying reliability"
author: "Brian Richardson"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output:
  html_document:
    code_folding: hide
---

```{r message = F, warning = F}

rm(list = ls())
library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggh4x)
library(kableExtra)
library(devtools)
load_all()

```

```{r warning = F}

# load simulation results from each of 10 clusters
setwd(dirname(getwd()))
sim.out.list <- lapply(
  X = 0:9,
  FUN = function(clust) {
    cbind(clust,
          read.csv(paste0("sim_data/sim6_ipw_reliability/sd",
                          clust, ".csv")))
  })

# combine simulation results into 1 data frame
sim.out <- bind_rows(sim.out.list)

# true MSM parameters
g <- c(2, 1.5, -1.5)

# number of sims per setting
n.rep <- nrow(sim.out) / n_distinct(dplyr::select(sim.out, n, B, rel))

# make long data frame
sim.out.long <- sim.out %>% 
  pivot_longer(cols = starts_with("ghat") |
                 starts_with("stde") |
                 starts_with("bste"), 
               names_to = "method.param",
               values_to = "val") %>% 
    mutate(method = factor(substr(method.param, 6, 7),
                           levels = c("NI", "CI"),
                           labels = c("Naive IPW",
                                      "IPW CS")),
           param = factor(substr(method.param, 9, 9)),
           name = factor(substr(method.param, 1, 4)),
           g.true = g[param]) %>% 
  dplyr::select(-method.param) %>% 
  group_by(clust, n, B, rel, method, param, name) %>% 
  mutate(id = row_number()) %>% 
  pivot_wider(names_from = name,
              values_from = val,
              id_cols = c(clust, n, B, rel, method,
                          param, g.true, id)) %>% 
  mutate(ci.lower = ghat - qnorm(0.975) * stde,
         ci.upper = ghat + qnorm(0.975) * stde,
         ci.cov = g.true >= ci.lower & g.true <= ci.upper,
         bcci.lower = ghat - qnorm(0.975) * bste,
         bcci.upper = ghat + qnorm(0.975) * bste,
         bcci.cov = g.true >= bcci.lower & g.true <= bcci.upper)

```

The goal of this program is to evaluate the IPW CS estimator with a nonlinear PS model and for a sequence of exposure reliability values.

## Setup

Consider the following setup:

* continuous confounder $L \sim U(0, 1)$,
* trivariate continuous exposure $\pmb{A} = (A_1, A_2, A_3)^T$ with $\pmb{A}|L$ having multivariate normal distribution $N_3\left(\begin{bmatrix} 0.4L^2 \\ -0.4L^2 \\ 0.2 - 0.1L^2 \end{bmatrix}, \begin{bmatrix} 0.09 & 0 & 0 \\ 0 & 0.09 & 0 \\ 0 & 0 & 0.09 \end{bmatrix}\right)$. 
* mismeasured exposure $\pmb{A}^* = (A^*_1, A^*_2,  A^*_3)^T = \pmb{A} + \pmb{\epsilon}$, where $\pmb{\epsilon} = (\epsilon_1, \epsilon_2, \epsilon_3)^T \sim N(\pmb{0}, \pmb{\Sigma}_e)$, $\pmb{\Sigma}_e = \begin{bmatrix} \sigma_e^2 & 0 & 0 \\ 0 & \sigma_e^2 & 0 \\ 0 & 0 & 0\end{bmatrix}$.
* binary outcome $Y$ with $\textrm{E}(Y|\pmb{A},L) =  \widetilde\gamma_0 + \widetilde\gamma_1 a_1 + \widetilde\gamma_2 a_2 + \widetilde\gamma_3 a_3 + \widetilde\gamma_4L + \widetilde\gamma_5 a_1L + \widetilde\gamma_6 a_2L + \widetilde\gamma_7 a_3L$, where $\widetilde{\pmb{\gamma}} = (\widetilde\gamma_0, \widetilde\gamma_1, \widetilde\gamma_2, \widetilde\gamma_3, \widetilde\gamma_4, \widetilde\gamma_5, \widetilde\gamma_6, \widetilde\gamma_7)^T = (0.4, 0.15, 0.15, 0.2, 0.1, 0.1, 0, -0.1)^T$.

This data generating process leads to the marginal structural model
$\textrm{E}\{Y(\pmb{a})\} = \gamma_0 + \gamma_1 a_1 + \gamma_2 a_2 + \gamma_3 a_3$, where $\pmb{\gamma} = (\gamma_0, \gamma_1, \gamma_2, \gamma_3)^T = (0.45, 0.2, 0.15, 0.15)^T$.

The estimand of interest in these simulations is the parameter $\pmb{\gamma}$ in the marginal structural model.

## Methods

In these simulations, we compare two methods:

1) Naive IPW
2) CS IPW

The naive method treats $A^*$ as $A$, assuming there is no measurement error. The CS method uses the Monte-Carlo corrected score versions of the estimating functions.

## Results

We first check for simulations with errors.

```{r message = F}

# summarize proportion of missing data by setting
sim.out.long %>% 
  group_by(method, n, B, rel) %>% 
  summarise(prop.error = mean(is.na(ghat))) %>% 
  filter(prop.error > 0)

```

```{r}

# extract simulation parameters
n <- unique(sim.out$n)
B <- unique(sim.out$B)
rel <- unique(sim.out$rel)

# make labels for plots
n.labs <- paste0("n = ", n)
names(n.labs) <- n

B.labs <- paste0("B = ", B)
names(B.labs) <- B

n.labs <- paste0("n = ", n)
names(n.labs) <- n

param.labs <- paste0("\u03b3", c("0", "1", "2", "3"))
names(param.labs) <- 1:4

rel.lab = unique(round(sim.out.long$rel, 2))
rel.lab[seq(2, n_distinct(rel.lab), 2)] <- ""

# colorblind friendly pallette
pal_light <- c('#EE6677', '#228833', '#4477AA', '#CCBB44', '#66CCEE', '#AA3377', '#BBBBBB')
pal_dark <- c('#991122', '#114419', '#223b55', '#6b611d', '#117799', '#55193b', '#5d5d5d')

```

```{r}

# create plots
make.plot <- function(title = F, param. = 2) {

  plot <- sim.out.long %>%
    filter(param == param.) %>% 
    mutate(rel = factor(rel)) %>% 
    ggplot(aes(x = rel, y = ghat, group = rel)) +
    geom_boxplot(width = 0.7, outlier.size = 0.5) +
    geom_hline(aes(yintercept = g.true),
               linetype = "dashed",
               color = "orange") +
    facet_grid(n ~ method,
               scales = "free", 
               labeller = labeller(n = n.labs)) +
    labs(y = "Parameter Estimate",
         x = "Exposure Reliability",
         fill = "Method",
         color = "Method") +
    theme_bw() +
    theme(legend.position = "bottom") +
    scale_x_discrete(labels = rel.lab)

  
  if (title) {
    plot <- plot +
      ggtitle(paste0("Empirical Distribution of Parameter Estimates"),
              subtitle = paste0(n.rep, " simulations per setting"))

  }
  return(plot)
}

```


```{r}

# create plots
make.plot.2 <- function(title = F, param. = 1) {
  
  plot <- sim.out.long %>% 
    filter(param == param.) %>% 
    pivot_longer(cols = ends_with(".cov"), 
                 names_to = "ci.type",
                 values_to = "ci.cov") %>% 
    mutate(ci.type = ifelse(ci.type == "bcci.cov",
                            "Bias-Corrected",
                            "Uncorrected")) %>% 
    group_by(rel, n, method, ci.type) %>% 
    summarise(cov = mean(ci.cov, na.rm = T)) %>% 
    ggplot(aes(x = rel,
               linetype = ci.type,
               y = cov)) +
    geom_line() +
    geom_hline(yintercept = 0.95,
               linetype = "dashed",
               color = "orange") +
    facet_grid(n ~ method,
               scales = "free", 
               labeller = labeller(n = n.labs)) +
    labs(y = "Coverage Probability",
         x = "Exposure Reliability",
         linetype = "CI Type") +
    theme_bw() +
    theme(legend.position = "bottom")
  
  if (title) {
    plot <- plot +
      ggtitle(paste0("Empirical Coverage Probabilities of 95% Confidence Intervals"),
              subtitle = paste0(n.rep, " simulations per setting"))

  }
  return(plot)
}

```

### Plots of Simulation Results

Below are boxplots of the empirical distributions of estimators for $\gamma_1$, and plots of coverage probabilities of confidence intervals.

```{r warning = F, message = F, eval = T}

make.plot()

# save plot
setwd(dirname(getwd()))
ggsave("sim_figures/sim6_ipw_reliability_1.png",
       width = 6,
       height = 4,
       dpi = 300)

```

```{r warning = F, message = F, eval = T}

make.plot.2()

# save plot
setwd(dirname(getwd()))
ggsave("sim_figures/sim6_ipw_reliability_2.png",
       width = 6,
       height = 4,
       dpi = 300)

```

