---
title: "Simulation 8: IPW with multiplicative measurement error"
author: "Brian Richardson"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output:
  html_document:
    code_folding: hide
---

```{r message = F, warning = F}

rm(list = ls())
library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggh4x)
library(kableExtra)
library(devtools)
load_all()

```

```{r warning = F}

# load simulation results from each of 10 clusters
setwd(dirname(getwd()))
sim.out.list <- lapply(
  X = 0:9,
  FUN = function(clust) {
    cbind(clust,
          read.csv(paste0("sim_data/sim8_ipw_multiplicative/sd",
                          clust, ".csv")))
  })

# combine simulation results into 1 data frame
sim.out <- bind_rows(sim.out.list)

# true MSM parameters
g <- c(0, 1, 1)

# number of sims per setting
n.rep <- nrow(sim.out) / n_distinct(dplyr::select(sim.out, n, B, rel))

# make long data frame
sim.out.long <- sim.out %>% 
  pivot_longer(cols = starts_with("ghat") |
                 starts_with("stde") |
                 starts_with("bste"), 
               names_to = "method.param",
               values_to = "val") %>% 
    mutate(method = factor(substr(method.param, 6, 7),
                           levels = c("NI", "CI"),
                           labels = c("Naive IPW",
                                      "IPW CS")),
           param = factor(substr(method.param, 9, 9)),
           name = factor(substr(method.param, 1, 4)),
           g.true = g[param],
           rel.lab = factor(rel,
                            labels = round(unique(rel), 2))) %>% 
  dplyr::select(-method.param) %>% 
  group_by(clust, n, B, rel, rel.lab, method, param, name) %>% 
  mutate(id = row_number()) %>% 
  pivot_wider(names_from = name,
              values_from = val,
              id_cols = c(clust, n, B, rel, rel.lab, method,
                          param, g.true, id)) %>% 
  mutate(ci.lower = ghat - qnorm(0.975) * stde,
         ci.upper = ghat + qnorm(0.975) * stde,
         ci.cov = g.true >= ci.lower & g.true <= ci.upper,
         bcci.lower = ghat - qnorm(0.975) * bste,
         bcci.upper = ghat + qnorm(0.975) * bste,
         bcci.cov = g.true >= bcci.lower & g.true <= bcci.upper)

```

The goal of this program is to evaluate the IPW CS estimator with a nonlinear PS model and under multiplicative measurement error.

## Setup

Consider the following setup:

* continuous confounder $L \sim N(0, 0.13)$,
* bivariate continuous exposure $\pmb{A} = (A_1, A_2)^T$ with $\pmb{A}|L$ having multivariate normal distribution $N_2\left(\begin{bmatrix} L^2 \\ -L^2 \end{bmatrix}, \begin{bmatrix} 1 & 0 \\ 0 & 1 \end{bmatrix}\right)$. 
* mismeasured exposure $\pmb{A^*} = (A_1^*, A_2^*)^T$ having multiplicative structure $A_i^* = \epsilon_iA_i$, where $\pmb{\epsilon} = (\epsilon_1, \epsilon_2) \sim N_2\left(\begin{bmatrix} 1 \\ 1 \end{bmatrix}, \begin{bmatrix} \sigma_e^2 & 0 \\ 0 & \sigma_e^2 \end{bmatrix}\right)$
* normal outcome $Y$ with $\textrm{E}(Y|\pmb{A},L) = \widetilde\gamma_0 + \widetilde\gamma_1 A_1 + \widetilde\gamma_2 A_2 + \widetilde\gamma_3 L$, where $\widetilde{\pmb{\gamma}} = (\widetilde\gamma_0, \widetilde\gamma_1, \widetilde\gamma_2, \widetilde\gamma_3)^T = (0, 1, 1, 1)^T$.

This data generating process leads to the marginal structural model
$\textrm{E}\{Y(\pmb{a})\} = \gamma_0 + \gamma_1 a_1 + \gamma_2 a_2$, where $\pmb{\gamma} = (\gamma_0, \gamma_1, \gamma_2)^T = (0, 1, 1)^T$.

The estimand of interest in these simulations is the parameter $\pmb{\gamma}$ in the marginal structural model.

The exposure reliability $rel  = \textrm{Var}(A_i) / \textrm{Var}(A_i^*)$ is varied within $[0.5, 1]$, for $i\in\{1,2\}$. Given $rel$, $\sigma^2_e$ is computed as $\textrm{Var}(A_i)*(rel^{-1} - 1) / (\textrm{Var}(A_i) + \textrm{E}(A_i)^2)$.

For these simulations, the measurement error is naively assumed to be additive, with covariance estimated using (for $k=10$)

$$\widehat{\pmb{\Sigma}}_e = \frac{\sum_{i=1}^{5} \sum_{j=1}^{k} (\pmb{A}^*_{ij} - \pmb{A}^*_{i.})^{T}(\pmb{A}^*_{ij} - \pmb{A}^*_{i.})}{\sum_{i=1}^{5}(k - 1)}.$$

## Methods

In these simulations, we compare two methods:

1) Naive IPW
2) CS IPW

The naive method treats $A^*$ as $A$, assuming there is no measurement error. The CS method uses the Monte-Carlo corrected score versions of the estimating functions.

## Results

We first check for simulations with errors.

```{r message = F}

# summarize proportion of missing data by setting
sim.out.long %>% 
  group_by(method, n, B, rel) %>% 
  summarise(prop.error = mean(is.na(ghat))) %>% 
  filter(prop.error > 0)

```

```{r}

# extract simulation parameters
n <- unique(sim.out$n)
B <- unique(sim.out$B)
rel <- unique(sim.out$rel)

# make labels for plots
n.labs <- paste0("n = ", n)
names(n.labs) <- n

B.labs <- paste0("B = ", B)
names(B.labs) <- B

n.labs <- paste0("n = ", n)
names(n.labs) <- n

param.labs <- paste0("\u03b3", c("0", "1", "2", "3"))
names(param.labs) <- 1:4

# colorblind friendly pallette
pal_light <- c('#EE6677', '#228833', '#4477AA', '#CCBB44', '#66CCEE', '#AA3377', '#BBBBBB')
pal_dark <- c('#991122', '#114419', '#223b55', '#6b611d', '#117799', '#55193b', '#5d5d5d')

```

```{r}

# create plots
make.plot <- function(title = F, param. = 2, y.lim = Inf) {
  
  # Calculate the counts of truncated values
  truncated_data <- sim.out.long %>%
    filter(param == param.) %>%
    group_by(rel.lab, n, method) %>%
    summarize(
      above = sum(ghat - g.true > y.lim),
      below = sum(ghat - g.true < -y.lim)
    ) %>%
    ungroup()
  
  # Merge the truncated counts with the original data
  sim.out.long.with.labels <- sim.out.long %>%
    filter(param == param.) %>%
    left_join(truncated_data, by = c("rel.lab", "n", "method")) %>%
    mutate(
      above.lab = ifelse(above == 0, "", as.character(above)),
      below.lab = ifelse(below == 0, "", as.character(below)))

  # Create the plot
  plot <- sim.out.long.with.labels %>%
    filter(abs(ghat - g.true) < y.lim) %>% 
    ggplot(aes(x = rel.lab)) +
    geom_boxplot(
      aes(y = ghat, group = as.factor(rel)),
      width = 0.7, outlier.size = 0.5) +
    geom_hline(
      aes(yintercept = g.true),
      linetype = "dashed", color = "orange") +
    geom_text(
      aes(y = g.true + y.lim + .3, label = above.lab),
      size = 3,
      color = "red"
    ) +
    geom_text(
      aes(y = g.true - y.lim - .3, label = below.lab),
      size = 3,
      hjust = 0.5,
      color = "red"
    ) +
    facet_grid(n ~ method,
               labeller = labeller(n = n.labs)) +
    labs(
      y = "Parameter Estimate",
      x = "Exposure Reliability",
      fill = "Method",
      color = "Method"
    ) +
    theme_bw() +
    theme(legend.position = "bottom")
  
  if (title) {
    plot <- plot +
      ggtitle(paste0("Empirical Distribution of Parameter Estimates"),
              subtitle = paste0(n.rep, " simulations per setting"))

  }
  return(plot)
}

```

```{r}

# create plots
make.plot.2 <- function(title = F, param. = 2) {
  
  plot <- sim.out.long %>% 
    filter(param == param.) %>% 
    pivot_longer(cols = ends_with(".cov"), 
                 names_to = "ci.type",
                 values_to = "ci.cov") %>% 
    mutate(ci.type = ifelse(ci.type == "bcci.cov",
                            "Bias-Corrected",
                            "Uncorrected")) %>% 
    group_by(rel, n, method, ci.type) %>% 
    summarise(cov = mean(ci.cov, na.rm = T)) %>% 
    ggplot(aes(x = rel,
               linetype = ci.type,
               y = cov)) +
    geom_line() +
    geom_hline(yintercept = 0.95,
               linetype = "dashed",
               color = "orange") +
    facet_grid(n ~ method,
               scales = "free", 
               labeller = labeller(n = n.labs)) +
    labs(y = "Coverage Probability",
         x = "Exposure Reliability",
         linetype = "CI Type") +
    theme_bw() +
    theme(legend.position = "bottom")
  
  if (title) {
    plot <- plot +
      ggtitle(paste0("Empirical Coverage Probabilities of 95% Confidence Intervals"),
              subtitle = paste0(n.rep, " simulations per setting"))

  }
  return(plot)
}

```

### Plots of Simulation Results

Below are boxplots of the empirical distributions of estimators for $\gamma_1$, and plots of coverage probabilities of confidence intervals.

```{r message = F}

make.plot(param. = 1, y.lim = Inf)
make.plot(param. = 2, y.lim = 1.5)
make.plot(param. = 3, y.lim = 1.5)

```

```{r message = F}

make.plot.2(param. = 1)
make.plot.2(param. = 2)
make.plot.2(param. = 3)

```

```{r warning = F, message = F, eval = T}

make.plot(param. = 2, y.lim = 1.5)

# save plot
setwd(dirname(getwd()))
ggsave("sim_figures/sim8_ipw_multiplicative_1.png",
       width = 7,
       height = 4,
       dpi = 300)

```

```{r warning = F, message = F, eval = T}

make.plot.2(param. = 2)

# save plot
setwd(dirname(getwd()))
ggsave("sim_figures/sim8_ipw_multiplicative_2.png",
       width = 7,
       height = 4,
       dpi = 300)

```

